Learning causal features with double regression

### 2019-05-16 paper/paper.tex
Sampling notation
abstract

\begin{abstract}
    To be written.
\end{abstract}

@@

More precisely, we measure, via an apparatus F, a multidimensional signal Y, that is a function of possible causes X, and other, unknown, factors that can be understood as noise N: $Y=f(X,N)$.
%
Of all possible cause (features of X), only a few account for Y, let them be represented as EX, with E a binary diagonal matrix: all zeroes except a few ones on the diagonal, which correspond to active causal features of X.
%
Let us assume that noise N is centred (adding, if need be, an intercept to the active causal features, to capture the bias in N), and additive (not a very strong assumption in the absence of further constraints on N).
%
Then, we can write our model of causal discovery as $Y=f(m(EX)+N)$, with f and m unknown functions.
%
Up to some rearrangement of the features of X, we can also suppose m linear.

@@

    \node[] (reg1) at (5,-3.25) {$\hat{E} = \overbrace{(X_2^\top X_2 + \lambda_X)^{-1} X_2^\top Y_2\underbrace{(Y_1^\top Y_1 + \lambda_Y)^{-1} Y_1^\top X_1}_{\text{1) } \hat{X} : \text{ regression from } Y \text{ to } X}}^{\text{2) regression from } X \text{ to } \hat{X}} $};
    \end{tikzpicture}
    \caption{Summary of our method blah blah blah}
    \label{fig:}
\end{figure}

>>>>

\begin{abstract}
Identify causes from observations is a core challenge to most empirical sciences. This endeavor is particularly challenging when (i) potential factors are difficult to manipulate separately and (ii) when observations are complex and/or multidimensional. To address this issue, we introduce in this paper the “back-and-forth regression” (BFR), a method that optimally identifies the most plausible causal factors from a set of covarying candidates. BFR consists in optimally (i) predicting the potential causes X from the observable Y, and subsequently (ii) identifying the only factors that contributed to those predictions. After detailing the proof of convergence, we show that BFR can systematically outperform least-square and related regression techniques, as well as cross-decomposition methods (e.g. canonical correlation analysis, partial least squares) on two distinct aspects: (i) causal identification and (ii) predictions outside independent and identically distributed datasets. We apply this method to empirical datasets and show that it retrieves plausible results. In particular, we apply BFR to neuroimaging recordings of 102 subjects presented with word sequences, and show, as expected, that the early brain responses appear to be specifically caused by low-level visual factors whereas late brain responses appear to be caused lexico-semantic factors, despite the fact that these two sets of factors are co-varying. We finish the paper by extending our approach to nonlinear cause-effects relationships. conclusion
\end{abstract}

@@

More precisely, we measure, via an apparatus F, a multidimensional signal Y, that is a function of possible causes X, and other, unknown, factors that can be understood as noise N: $Y=f(X,N)$.
%
Of all possible causes (features of X), only a few account for Y, let them be represented as EX, with E a binary diagonal matrix: all zeroes except a few ones on the diagonal, which correspond to active causal features of X.
%
Let us assume that noise N is centred (adding, if need be, an intercept to the active causal features, to capture the bias in N), and additive (not a very strong assumption in the absence of further constraints on N).
%
Then, we can write our model of causal discovery as $Y=f(m(EX)+N)$, with f and m unknown functions.
%
Up to some rearrangement of the features of X, we can also suppose m linear.

@@

    \node[] (sim1) at (0,-3.25) {\scalebox{0.85}{$X \sim P(X)$}};
    \node[] (sim2) at (0,-3.65) {\scalebox{0.85}{$N \sim P(N)$}};
    \node[] (reg1) at (5,-3.25) {$\hat{E} = \overbrace{(X_2^\top X_2 + \lambda_X)^{-1} X_2^\top Y_2\underbrace{(Y_1^\top Y_1 + \lambda_Y)^{-1} Y_1^\top X_1}_{\text{1) } \hat{X} : \text{ regression from } Y \text{ to } X}}^{\text{2) regression from } X \text{ to } \hat{X}} $};
    \end{tikzpicture}
    \caption{Summary of our method blah blah blah}
    \label{fig:}
\end{figure}
